---
title: "BIOE 215 Week6"
format: html
editor: visual
---

## browser()

Browser lets you see the world as the function sees it. When you call browser () it allows you to pause the execution of a function so that you can poke around in it.

```{r}
library(palmerpenguins)
library(tidyverse)

traits_r2 <- function(trait1, trait2) {
  #browser()
  summary(lm(trait1 ~ trait2))$r.squared
}

penguins %>% 
  summarise(flipper_bill_r2 = traits_r2
            (flipper_length_mm, bill_length_mm))
```

**Q1:** How would you describe the error? I'm not asking you describe the *cause* of the error yet. Describe how the output is different than what you would expect.

Well we have an empty data frame. We expected to get a tibble with the r\^2 value for the lm of trait1 and trait 2.

**Q2:** Where would you add `browser()` to explore the cause of the error?

I would put browser() after I define my function, after the first open curly bracket. This is because browser() is a function and needs inputs.

**Q3:** Does the body of `traits_r2` use list or atomic vector indexing? Does it use indexing by position, logic, or name?

Summary is followed by a \$ it is list indexing by name.

**Q4:** What's the cause of the error? How would you fix it?

Yes there's a typo. There's a . missing in rsquared. It should be r.squared. Run this in browser: names(summary(lm(trait1 \~ trait2)))

## Adding `group_by()`

```{r}
# Pipeline 1
penguins %>% 
  group_by(species) %>% 
  summarize(flipper_bill_r2 = traits_r2(flipper_length_mm, bill_length_mm))

# Pipeline 2
penguins %>% 
  group_by(species, island) %>% 
  summarize(flipper_bill_r2 = traits_r2(flipper_length_mm, bill_length_mm),
            .groups = "drop")

# Pipeline 3
penguins %>% 
  group_by(species, island, sex) %>% 
  summarize(flipper_bill_r2 = traits_r2(flipper_length_mm, bill_length_mm),
            .groups = "drop")
```

**Q5:** How many times does Pipeline 1 call `traits_r2()`? How about Pipeline 2?

Pipeline 1: is called once per species so 3 times

Pipeline 2: is called once per species AND island, each unique member of every group in `group_by()`. So it is getting called 5 times in pipeline2.

**Q6:** Create Pipeline 3 that additionally groups by sex. How many times does Pipeline 3 call `traits_r2()`?

13 times!

## group_by-summarize w error

```{r}
#| eval: false
set.seed(12345)
penguins2 <- penguins %>% 
  drop_na(sex) %>% 
  sample_n(25) 
penguins2[7, 3:6] <- NA

traits_r2 <- function(trait1, trait2) {
  browser()
  summary(lm(trait1 ~ trait2))$r.squared
}

penguins2 %>% 
  group_by(species, island, sex) %>% 
  summarize(flipper_bill_r2 = traits_r2(flipper_length_mm, bill_length_mm),
            .groups = "drop")
```

**Q7:** How would you describe the error?

I got an error and it stopped. I was expecting a similar tibble to pipeline 3 above with about 13 rows of data grouped by species, island, and sex with the r-squared value for each entry.

**Q8:** Use `browser()` to diagnose the error. Hint: `c` will tell the debugger to continue until the next time it's called.

Group 6 does not exist.

**Q9:** How would you fix the error?

drop_na

# **Visualizing data with `ggplot()`**

`ggplot()` has four major components:

-   Data

-   Scales

-   Geometries

-   Aesthetics

```{r}
library(palmerpenguins)
library(tidyverse)

ggplot(data = penguins, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point()

head(penguins)
```

Change the aesthetic to show the relationship with bill_depth_mm and set the color to be by species.

```{r}
ggplot(data = penguins, aes(x = body_mass_g, y = bill_depth_mm, color = species)) +
  geom_point(aes(color = species))
```

## Customizing scales

```{r}
ggplot(data = penguins) +
  # You can also define aesthetics directly inside the geometries
  geom_point(aes(x = body_mass_g, 
                 y = flipper_length_mm,
                 color = species)) +
  # x aesthetic, continuous scale
  scale_x_continuous(
    # change the axis name
    name = "Body mass (g)",
    # change the limits
    limits = c(2000, 8000)
  ) +
  # color aesthetic, manual scale
  scale_color_manual(
    # set the values for the colors
    values = c(Adelie = "cornflowerblue",
               Chinstrap = "firebrick",
               Gentoo = "darkorchid")
  )
```

**Q12:** What function would you use to customize the y-axis? Use that function to expand the y-axis limits to include 150 and 250.

**Q13:** Look up the help for `scale_color_brewer()`. Change the color scale to use the \"Dark2\" color palette.

```{r}
#From the plot above...
ggplot(data = penguins) +
  # You can also define aesthetics directly inside the geometries
  geom_point(aes(x = body_mass_g, 
                 y = flipper_length_mm,
                 color = species)) +
  # x aesthetic, continuous scale
  scale_x_continuous(
    # change the axis name
    name = "Body mass (g)",
    # change the limits
    limits = c(2000, 8000)
  ) +
  # changing the y-axis
  scale_y_continuous(
    # change the axis name
    name = "Flipper length (mm)",
    # change the limits
    limits = c(150, 250)
  ) + 
  # color aesthetic, manual scale
  scale_color_brewer(palette = "Dark2")
```
